---
title: "Homework 2: Bike Sharing Data Visualization"
author: "YOUR NAME"
output: prettydoc::html_pretty
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)

# Disable a warning message.
options(dplyr.summarise.inform = FALSE)
```

*Your introduction here*

## Read Data

Read the data from the CSV file.

```{r read-data, message=FALSE}
daily_rides <- read_csv("data/day.csv")
```

The following are two different ways of getting a quick look at the data. (Which one looks better in a report? Which one is more useful to a programmer?)

```{r glimpse-data}
daily_rides %>% glimpse()
```


```{r head-data}
daily_rides %>% head() %>% kable()
```

## Exercise 1: Label days of week

*Replace this line with your explanation of how you determined which number corresponds to which day.* Remove the asterisks so it isn't italic in your report.

```{r label-weekdays}
daily_rides <- daily_rides %>%
  mutate(day_of_week = factor(day_of_week, levels = c(0, 1, 2, 3, 4, 5, 6), labels = c("DAY", "ANOTHER DAY", "...", "...", "...", "...", "...")))
```

## Exercise 2: Describe a row

*Your description here* without the asterisks.

## Exercise 3: Rides by date, by rider type

```{r rides-by-date-membership}
# your code here (remove this line)
```

*Your interpretation here*

## Exercise 4: Mapping vs Faceting

```{r rides-membership-workingday-1}
# your code here (remove this line)
```


```{r rides-membership-workingday-2}
# your code here (remove this line)
```

*Your comparison and contrast here*

## Exercise 5: How does ridership vary over a typical week?

```{r ride-distributions-by-weekday}
# your code here (remove this line)
```

*Your description of what the plot tells you about the data here*


## Exercise 6: Plot of your choice

Add your plot code here.

*Your description of what the plot tells you about the data here*
