---
title: "HW 02 General feedback"
output: 
  tufte::tufte_html:
    css: ../hw.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
options(dplyr.summarise.inform = FALSE)
```

Here were some things we noticed commonly going wrong in hw2.

```{r read-data, message=FALSE, include=FALSE}
hourly_rides <- read_csv("../../../src/hw/hw02-vis/data/hour.csv")
```

# Exercise 1: Label days of week

Many people didn't include descriptions of how they figured out that these were the correct labels.

"I looked at a calendar and January 1 2011 was a Saturday, which is coded as 6" would have been acceptable.

```{r label-weekdays, include=FALSE}
hourly_rides <- hourly_rides %>%
  mutate(day_of_week = factor(day_of_week, levels = c(0, 1, 2, 3, 4, 5, 6), labels = c("Sun", "Mon", "Tue", "Wed", "Thur", "Fri", "Sat")))
```


# Exercise 5

Note the difference between specifying an attribute:

```{r overplot-attr}
ggplot(hourly_rides, aes(x = temp, y = rides)) +
  geom_point(color = "blue") +
  facet_wrap(~ member_type) + 
  labs(x = "Normalized temperature", y = "Rides")
```

... and mapping an aesthetic:

```{r overplot-aes}
ggplot(hourly_rides, aes(x = temp, y = rides)) +
  geom_point(aes(color = "blue")) +
  facet_wrap(~ member_type) + 
  labs(x = "Normalized temperature", y = "Rides")
```
