---
title: "Predictive Modeling"
author: "K Arnold"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    self_contained: FALSE
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
source("../slides-common.R")
slideSetup()
knitr::opts_chunk$set(echo = TRUE)
```

class: center, middle

# A Whimsical Take on AI

[slides](https://docs.google.com/presentation/d/1oWLJmM9piLf4dGbmwQj5heRAnLulR98OSUZNXQyHQRI/edit#slide=id.g5efabcf039_0_67)

<br>

.small[
[An Ethics of Artificial Intelligence Curriculum for Middle School Students](https://docs.google.com/document/d/1e9wx9oBg7CR0s5O7YnYHVmX7H7pnITfoDxNdrSGkp60/edit#heading=h.lj79p5m41lr3)
]

---

class: middle

## Bingo Activity

* In *cohorts* (groups of 4 or 5)...
* *Teams* -> *your cohort's channel* -> *Files* tab -> "Prediction Bingo Activity"


---

## Predictive Modeling

* A powerful tool to turn data into action.
* It works because God made the universe predictable (and successful prediction rewarding)
* **Need for wisdom**: It can be used for great good and great harm

---

## Power of Predictive Modeling

* **Medicine**: wearable monitor for seizures or falls, detect malaria from blood smears, find effective drug regimens from medical records
* **Drug Discovery**: predict the efficacy of a synthesis plan for a drug
* **Precision Agriculture**: predict effect of micro-climate on plant growth
* **Urban Planning**: forecast resource needs, extreme weather risks, ...
* **Government**: classify feedback from constituents
* **Retail**: predict items in a grocery order
* **Recommendation systems**: Amazon, Netflix, YouTube, ...
* **User interfaces**: gesture typing, autocomplete / autocorrect

and so much more...

---

## The universe is surprisingly predictable

* God created the world with actionable structure
  * We gradually learn how to perceive that structure and act within it.
  * The better our perceptions align with how the universe is structured, the better our actions
<!--* Every living thing does!  All living things take in information, make inferences about the world from it, and use those inferences to inform actions.-->
  * We can discover that structure by learning to be less surprised by what we see ( = predicting our perceptions)
* Perceptions are thus both accurate and fallable.

---

## Predictive modeling technology: Need for wisdom

* Potential for great good
* But also great harm:
  * Lack of **fairness** in facial recognition, sentencing, lending, job applicant scoring, ...
  * Lack of **transparency** in how "Big Data" systems make conclusions
  * Lack of **privacy** as data is increasingly collected and aggregated
  * Amplification of extreme positions in social media, YouTube, etc.
  * Oversimplification of human experience
  * Hidden human labor
  * Illusion of objectivity
  * ...!

---

class: middle, center

# Stating and refining the question

---

# Six types of questions

1. **Descriptive:** summarize a characteristic of a set of data
1. **Exploratory:** analyze to see if there are patterns, trends, or relationships between variables (hypothesis generating)
1. **Inferential:** analyze patterns, trends, or relationships in representative data from a population
1. **Predictive:** make predictions for individuals or groups of individuals
1. **Causal:** whether changing one factor will change another factor, on average, in a population
1. **Mechanistic:** explore "how" as opposed to whether

.footnote[
Leek, Jeffery T., and Roger D. Peng. "What is the question?." Science 347.6228 (2015): 1314-1315.
]

---

# Ex: Viral illnesses

1. **Descriptive:** severity of viral illnesses in a set of data collected from a group of individuals
--

1. **Exploratory:** examine relationships between a range of dietary factors and viral illnesses
--

1. **Inferential:** examine whether any relationship between dietary factors and viral illnesses found in the sample hold for the population at large

--
1. **Predictive:** given a person's demographics and diet, predict the severity of illness

--
1. **Causal:** whether people who were randomly assigned to eat a diet high in fresh fruits and vegetables or one that was low in fresh fruits and vegetables contract more severe viral illnesses

--
1. **Mechanistic:** how a diet high in fresh fruits and vegetables leads to a reduction in the severity of viral illnesses

---

## Our focus: Prediction

* **Why** are we doing this?
  * Predictions enable decision-making
  * trying to predict helps us understand.
* **What** are we doing?
  * Predict something unknown from something known. Specifically: complete-the-table model
  * Mostly we'll *assume independent observations* (e.g., generalize across people)
  * Sometimes we'll *forecast* (e.g., predict how a time series will continue)
* **How**: We’ll look at...
  * methods that consider similar examples (Nearest Neighbors)
  * methods that look at overall trends (linear/logistic regression)
  * more advanced methods, time permitting

---

## Example: Home Sale Prices

```{r include=FALSE}
ames_home_sales <- AmesHousing::make_ames()
```

From Ames, Iowa home sales, 2006-2010. (De Cock, 2011)

.small[
```{r echo=FALSE}
ames_home_sales %>%
  select(Lot_Area, Total_Bsmt_SF, Gr_Liv_Area, Garage_Cars, Sale_Price) %>% 
  head(5) %>% 
  kableExtra::kbl() %>% 
    kableExtra::column_spec(5, background = "#8C2131", color = "white")
```
]
(`r nrow(ames_home_sales)` total rows)

* *Y*: *response* variable (aka *outcome*, *dependent variable*): `Sale_Price`
* *X*: *features* (aka *predictors*, *covariates*, etc.): everything else

Note: `X` is much easier to measure than `Y`
