---
title: "UN Votes"
author: "Mine Çetinkaya-Rundel, adapted by Kenneth Arnold"
date: "`r Sys.Date()`"
output: 
  learnr::tutorial:
    progressive: false
    allow_skip: true
    toc: false
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
```

## Introduction

Let's take a look at the voting history of countries in the United Nations 
General Assembly. We will be using data from the `unvotes` package. 
Additionally, we will make use of the `tidyverse` and `lubridate` packages 
for the analysis, and the `DT` package for interactive display of tabular 
output.

```{r load-packages, warning=FALSE, message=FALSE}
library(unvotes)
library(tidyverse)
library(lubridate)
library(DT)
theme_set(theme_bw())
```

The `unvotes` package provides three datasets we can work with: 

1. `un_roll_calls`: Each roll call vote of the United Nations General Assembly
2. `un_roll_call_issues`: Topic classifications of roll call votes of the United Nations General Assembly
3. `un_votes`: Information on the voting history of the United Nations General Assembly

Each of these datasets contains a variable called `rcid` (roll call id) which 
can be used as a unique identifier to join them with each other.

```{r}
un_votes %>% head(10)
un_roll_calls %>% head(10)
un_roll_call_issues %>% head(10)
```


## Wrangle and Visualize

Let's create a data visualization that displays how the voting record of the 
US changed over time on a variety of issues, and compares it to two other 
countries: the United Kingdom and Turkey.

```{r plot-yearly-yes-issue, fig.width=10, warning=FALSE, exercise=TRUE}
un_votes %>%
  mutate(country =
           case_when(
             country == "United Kingdom of Great Britain and Northern Ireland" ~ "UK & NI",
             country == "United States of America"                             ~ "US",
             TRUE                                                              ~ country
             )) %>%
  filter(country %in% c("UK & NI", "US", "Turkey")) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  inner_join(un_roll_call_issues, by = "rcid") %>%
  group_by(country, year = year(date), issue) %>%
  summarize(
    votes = n(),
    percent_yes = mean(vote == "yes"),
    .groups = 'keep'
    ) %>%
  filter(votes > 5) %>%  # only use records where there are more than 5 votes
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_point() +
    geom_smooth(method = "loess", formula = y ~ x, se = FALSE) +
    facet_wrap(~ issue) +
    labs(
      title = "Percentage of 'Yes' votes in the UN General Assembly",
      subtitle = "1946 to 2015",
      y = "% Yes",
      x = "Year",
      color = "Country"
    ) +
  theme(legend.position = "bottom")
```

We can easily change which countries are being plotted by changing which 
countries the code above `filter`s for. Note that the country name should be 
spelled and capitalized exactly the same way as it appears in the data. See 
the [Appendix](#appendix) for a list of the countries in the data.

## References

1. David Robinson (2017). unvotes: United Nations General Assembly Voting Data. 
R package version 0.2.0. https://CRAN.R-project.org/package=unvotes.
2. Erik Voeten "Data and Analyses of Voting in the UN General Assembly" 
Routledge Handbook of International Organization, edited by Bob Reinalda 
(published May 27, 2013).
3. Much of the analysis has been modeled on the examples presented in the 
[unvotes package vignette](https://cran.r-project.org/web/packages/unvotes/vignettes/unvotes.html).

## Appendix {#appendix}

Below is a list of countries in the dataset:

```{r list-countries}
un_votes %>% 
  select(country) %>%
  arrange(country) %>% 
  distinct() %>%
  datatable()
```
