---
title: 'Homework 10: Predictive Modeling'
author: "YOUR NAME"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
theme_set(theme_bw())
options(dplyr.summarise.inform = FALSE) # silence a warning message
```


```{r utilities, echo=FALSE}
add_predictions <- function(data, ...) {
  purrr::imap_dfr(
    rlang::dots_list(..., .named = TRUE),
    function(model, model_name) {
      model %>%
        predict(data) %>%
        bind_cols(data) %>%
        mutate(model = !!model_name)
    }
  )
}
```

## Getting Started

### Load data

```{r load-data, include=TRUE}
daily_rides <- read_csv("data/day-hw10.csv", col_types = cols_only(
  date = col_date(),
  year = col_integer(),
  workingday = col_character(),
  temp = col_double(),
  atemp = col_double(),
  casual = col_double(),
  registered = col_double()
)) %>% mutate(across(c(workingday, year), as_factor))
```


### Exploratory Analytics

### Train-test split

## Linear Regression using Temperature

### Look inside the model

### Predictions

### Residuals Histogram

### Observed by Predicted

### Validate the model on the test set

#### Predictions

#### Residuals

#### Observed by Predicted

### Quantify errors

#### Summarize

## Linear Regression using Temperature and Working Day

## Decision Tree Regression

## Wrap-up
