---
title: "Quiz 7 Supplement"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#remotes::install_github("rundel/learnrhash")
library(learnr)
library(tidyverse)
library(lubridate)
# Disable a warning message.
options(dplyr.summarise.inform = FALSE)

library(modeldata)
data("ames", package = "modeldata")
```

## Visualization

### Reproduce a plot

Write code to reproduce the following visualization, intended to show
how home sale price (`Sale_Price`) varies with the number of bathrooms.
Use the `Full_Bath` and `Half_Bath` columns of the `ames` data frame, and count
a half bath as 0.5 bathrooms.

```{r bathroom-plot-bad, echo=FALSE}
ames %>% 
  mutate(baths = Full_Bath + .5 * Half_Bath) %>% 
  ggplot(aes(x = baths, y = Sale_Price)) +
    geom_point()
```

```{r bathroom-plot-reproduce, exercise=TRUE}

```

## Data Wrangling

In which neighborhood (`Neighborhood` column) was the home with the
highest `Sale_Price`?

```{r most-expensive-neighborhood, exercise=TRUE}
ames %>% ...
```


```{r most-expensive-neighborhood-sol, include=FALSE}
ames %>%
  arrange(desc(Sale_Price)) %>% 
  select(Neighborhood, Sale_Price)
```

Which neighborhood has the lowest median `Sale_Price`?

```{r lowest-median-neighborhood, exercise=TRUE}

```


```{r lowest-median-neighborhood-sol, include=FALSE}
ames %>%
  group_by(Neighborhood) %>%
  summarize(m = median(Sale_Price)) %>% 
  arrange(m)
```
<!--


## Submit

```{r context="server"}
learnrhash::encoder_logic()
```

If you have completed this quiz and are happy with all of your solutions,
please *click the button below* to generate (or update) your hash. Then,
copy-paste the entire hash into Moodle.

```{r encode, echo=FALSE}
shiny::tags$div(
  shiny::fixedRow(shiny::column(width = 3, 
                                shiny::actionButton("hash_generate", "Generate Submission")), 
                  shiny::column(width = 7), shiny::column(width = 2)), 
  shiny::tags$br(), learnrhash:::wrapped_verbatim_text_output("hash_output", TRUE),
  shiny::tags$br())
```
-->
