---
title: "Lab 02: Global Plastic Waste"
output: github_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

## The Data
```{r load_data, include=FALSE, echo=FALSE, warning=FALSE}
plastic_waste <- read_csv("data/plastic-waste.csv")
```

## Exercise 1
```{r plastic_waste_perCapita, fig.align='center', fig.width=8, fig.height=6, warning=FALSE}
plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap, fill = continent)) +
  geom_histogram(bins = 50) +
  facet_wrap(~continent) +
  labs(x = "Plastic Waste per Capita", y = "Count", fill = "Continent")

plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap, color = continent, fill = continent)) +
  geom_density(alpha = 0.07) +
  labs(x = "Plastic Waste per Capita", y = "Count")
```

* According to these plots, most of the countries do **not** surpass 1 kilogram of plastic waste per day, except for Trinidad and Tobago. 

## Exercise 2

```{r lower_alpha, fig.align='center', fig.width=8, fig.height=6, warning=FALSE}
plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap, color = continent, fill = continent)) +
  geom_density(alpha = 0.15) +
  labs(x = "Plastic Waste per Capita", y = "Count")
```

## Exercise 3

* 

## Exercise 4

```{r violin_plot, fig.align='center', fig.width=8, fig.height=6, warning=FALSE}
plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap, y = continent)) +
  geom_boxplot() +
  labs(x = "Plastic Waste per Capita", y = "Continent")

plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap, y = continent)) +
  geom_violin() +
  labs(x = "Plastic Waste per Capita", y = "Continent")
```

* A violin plot displays a continuous distribution. It is displayed similarily as a boxplot but it is displayed as a mirrored density plot. A violin plot shows where there is a high density of points. In other words, a violint plot allows us to see the shape of the distribution. While we can draw quantiles in a violin plot, a box plot always displays the interquartile range, the median, and the outliers. 

## Exercise 5

```{r mismanaged_waste, fig.align='center', fig.width=8, fig.height=6, warning=FALSE}
plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste_per_cap)) +
  geom_point() +
  labs(x = "Plastic Waste per Capita", y = "Mismanaged Plastic Waste per Capita")
```

* There seems to a slight positive relationship between `plastic waste per capita` and `mismanaged plastic waste per capita`.

## Exercise 6

```{r mismanaged_waste_continent, fig.align='center', fig.width=8, fig.height=6, warning=FALSE}
plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste_per_cap, color = continent)) +
  geom_point() +
  labs(x = "Plastic Waste per Capita", y = "Mismanaged Plastic Waste per Capita", color = "Continent")
```

* There doesn't seem to be any clear distinctions between continents with respect to how `plastic waste per capita` and `mismanaged plastic waster per capita` are associated. Most of the outliers seem to be countries from Africa, Asia, and Oceania, but most of the countries are clustered. 

## Exercise 7

```{r total_population, fig.align='center', fig.width=8, fig.height=6, warning=FALSE}
plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap, y = total_pop, color = continent)) +
  geom_point() +
  labs(x = "Plastic Waste per Capita", y = "Total Population", color = "Continent")

plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap, y = coastal_pop, color = continent)) +
  geom_point() +
  labs(x = "Plastic Waste per Capita", y = "Coastal Population", color = "Continent")
```

* There doesn't seem to be a strong linear association between `plastic waste per capita` and `total population` and/or `coastal population`. One conclusion that could be made is the fact that *Asia* seem to have countries with larger `total population` and `coastal population`, which makes sense (China, India, Japan, etc.), but do not produce the larget amount of plastic waste per capita. Instead, countries with smaller populations seem to produce a slightly higher snumber of plastic waste per capita. While there might be several reasons to why that is the case, one hypothesis might be if a country produces the same amount of plastic waste but has a smaller population, then it will have a slightly higher number than a country that produces the same amount of plastic waste but with a bigger population. 

* *Note*: It would be interesting to study how and why countries with small populations produce a significant amount of plastic waste.

## Exercise 8
```{r, fig.align='center', fig.width=8, fig.height=6, warning=FALSE}
plastic_waste %>% 
  mutate(coastal_pop_prop = coastal_pop / total_pop) %>%
  filter(plastic_waste_per_cap < 3) %>%
  ggplot(aes(x = coastal_pop_prop, y = plastic_waste_per_cap, color = continent)) + 
    geom_point() +
    geom_smooth(color = "black") +
    scale_color_viridis_d() +
    labs(x = "Coastal Population Proportion (Coastal / Total population)", 
         y = "Plastic Waste per Capita ", 
         color = "Continent",
         title = "Plastic Waste vs. Coastal Population Proportion",
         subtitle = "by Continent") +
    theme_minimal()
```
