---
title: "Lab 07: Making Predictive Models"
author: "TEAM"
date: "DATE"
output: github_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
 options(scipen = 5) # encourage metrics to print in fixed-point notation
```


## Data

```{r load-and-subset-data}
data(ames, package = "modeldata")
ames <- ames %>% 
  filter(Gr_Liv_Area < 4000, Sale_Condition == "Normal")
```

The data contains sale prices for ____ homes.

## Exploratory analysis


## Modeling

### Hold out some data to use for validation

```{r train-test-split}
# Seed the random number generator
set.seed(10)
# Split our data randomly
ames_split <- initial_split(ames, prop = 2/3)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

*Your narrative answers here*

### Specify the model

```{r model-spec}
my_model_spec <- 
  parsnip::linear_reg() %>% 
  set_mode("regression") %>% 
  set_engine("lm")
```


### Fit the model on data

```{r train-model-1}
# code here
```

*Your narrative answers here*

### How well did it do?

```{r predict-model-1}
# code here
```


*Your narrative answers here*

On average, this model's predictions were off by about $___ on its training data.


### How well did it do... on *unseen* data?


```{r eval-model-1}
# code
```

### Is that good?

On average, this model's predictions were off by about $___.

*Your narrative answers here.*

## More features

```{r}
# code
```

*Your narrative answers here.*


## LOTS more features!

```{r}
# code
```


*Your narrative answers here.*
